---
title: "Chesapeake Bay Watershed and LULC Analysis"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
---

The goal of this analysis is to delineate watersheds for points associated with the CBW-LULC project. This script relies on both the spatial capabilities of R (Raster, sp, and RGDAL packages) in addition to the hydrologic analysis tools in [WhiteBox Tools](http://www.uoguelph.ca/~hydrogeo/WhiteboxTools/index.html) scripting software. Required datasets inlcude a [NHDplus 30m DEM](http://www.horizon-systems.com/NHDPlus/NHDPlusV1_02.php), [the gagesII database](https://water.usgs.gov/GIS/metadata/usgswrd/XML/gagesII_Sept2011.xml), and finally, [NLCD-LULC data](https://water.usgs.gov/GIS/metadata/usgswrd/XML/gagesII_Sept2011.xml) from the USGS. All code is house at [GitHub](https://github.com/FloodHydrology/InundationHydRology)

For more on wateshed delineation in WhiteBox GAT and WhiteBoxTools, [see this block post from Jon Lindsay](https://whiteboxgeospatial.wordpress.com/2014/05/04/mapping-watersheds-in-whitebox-gat/).

#Step 1: Data Organization
Note, data is currently housed in Nate's SESYNC workspace.  He will eventually migrate this to the Palmer Lab folder. 

```{r, message=FALSE, warning=FALSE, results='hide'}
#Clear Memory
rm(list=ls(all=TRUE))

#Define relevant working directories
data_dir<-"//storage.research.sesync.org/njones-data/Research Projects/LULC_CBW/Spatial_Data/"
scratch_dir<-"C:\\ScratchWorkspace/"
wbt_dir<-"C:\\WBT/whitebox_tools"

#Add appropriate libararies
library(sf)
library(raster)
library(rgdal)
library(rgeos)
library(dplyr)
library(devtools)

#Download package from GIT
install_github("FloodHydrology/InundationHydRology")
library(InundationHydRology)

#Download required data
dem<-raster(paste0(data_dir,"NHDPlus02/Elev_Unit_a/elev_cm"))
  p<-dem@crs
HUC08<-st_read(paste0(data_dir,"NHDPlus02/Subbasin.shp"))
  HUC08<-st_transform(HUC08, crs=paste(p))
```

Step 2: Delineate watershed using the LULC_WatershedAnalysis r_script in the InundationHydRology package. 

```{r}
#Define pnts
pnts<-read.csv(paste0(data_dir,"input.csv"))
pnts<-st_as_sf(pnts, 
               coords=c("Longitude","Latitude"), 
               crs="+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")
pnts<-st_transform(pnts, paste(dem@crs))

#Select Subbasin
mask<-HUC08
mask<-mask[pnts,]

fun<-function(n){
  tryCatch(WatershedAnalysis(  
                dem=dem,  
                pnts=pnts,
                unique_id = "FldActI",
                mask=mask[n,], 
                threshold=300,
                scratch_dir=scratch_dir,
                data_dir = data_dir,
                wbt_dir = wbt_dir,
                output_dir = paste0(data_dir,"watershed/")), 
           error=function(e) c(n, rep(0,24)))}

#Run function
lapply(seq(1,length(mask)), fun)
```

So, I've encourntered problems at i=23. I'm not sure why.  Next step will be to investigate this. 

