---
title: "Chesapeake Bay Watershed and LULC Analysis"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
---

The goal of this notebook is to demonstrate the watershed delineation technique we plan to employ in the CBW-LULC anlaysis. This computation relies on both the spatial capabilities of R (Raster, sp, and RGDAL packages) in addition to the hydrologic analysis tools in [WhiteBox Tools](http://www.uoguelph.ca/~hydrogeo/WhiteboxTools/index.html) scripting software. Required dataset inlcude a [NHDplus 30-m DEM](http://www.horizon-systems.com/NHDPlus/NHDPlusV1_02.php), [the gagesII database](https://water.usgs.gov/GIS/metadata/usgswrd/XML/gagesII_Sept2011.xml), and finally, [NLCD-LULC data](https://water.usgs.gov/GIS/metadata/usgswrd/XML/gagesII_Sept2011.xml) from the USGS.

For more on wateshed delineation in WhiteBox GAT and WhiteBoxTools, [see this block post from Jon Lindsay](https://whiteboxgeospatial.wordpress.com/2014/05/04/mapping-watersheds-in-whitebox-gat/).

#Step 1: Data Organization
Note, data is currently housed in Nate's SESYNC workspace.  He will eventually migrate this to the Palmer Lab folder. 

```{r, message=FALSE, warning=FALSE, results='hide'}
#Clear Memory
rm(list=ls(all=TRUE))

#Define relevant working directories
data_dir<-"//storage.research.sesync.org/njones-data/Research Projects/LULC_CBW/Spatial_Data/"
scratch_dir<-"C:\\ScratchWorkspace\\"
wbt_dir<-"C:/WBT/whitebox_tools"

#Add appropriate libararies
library(sp)
library(raster)
library(rgdal)
library(rgeos)
library(dplyr)

#Download package from GIT
library(devtools)
install_github("FloodHydrology/InundationHydRology")
library(InundationHydRology)

#Download required data
dem<-raster(paste0(data_dir,"NHDPlus02/Elev_Unit_a/elev_cm"))
  p<-dem@crs
gages<-readOGR(paste0(data_dir,"NHDPlus02/StreamGageEvent.shp"))
  gages<-gages[gages$SUBREGION %in% c("0205","0206","0207","0208"),]
  gages<-spTransform(gages, p)
HUC08<-readOGR(paste0(data_dir,"NHDPlus02/Subbasin.shp"))
  HUC08<-spTransform(HUC08, p)
LULC2001<-raster(paste0(data_dir,"nlcd_2001_landcover_2011_edition_2014_10_10/nlcd_2001_landcover_2011_edition_2014_10_10/nlcd_2001_landcover_2011_edition_2014_10_10.img"))
LULC2006<-raster(paste0(data_dir,"nlcd_2006_landcover_2011_edition_2014_10_10/nlcd_2006_landcover_2011_edition_2014_10_10.img"))
LULC2011<-raster(paste0(data_dir,"nlcd_2011_landcover_2011_edition_2014_10_10/nlcd_2011_landcover_2011_edition_2014_10_10.img"))

```

#Step 2: Prep Data!
Here we need to (1) calculate flow direction raster and (2) convert the stream into a raster. 

```{r, message=FALSE, warning=FALSE, results='hide'}
LULC_WatershedAnalysis<-function(
  dem=dem,  
  pnts=pnts,
  unique_id="SOURCE_FEA",
  streams=streams,
  mask=mask, 
  threshold=1000,
  LULC2001=LULC2001,
  LULC2006=LULC2006,
  LULC2011=LULC2011, 
  data_dir=data_dir){

#Mask dem and gages
dem<-crop(dem, mask)
pnts<-pnts[mask,]

#Add UID to pnts
pnts$ID<-seq(1,length(pnts))

#Export DEM and stream layer to local working directory
writeRaster(dem, paste0(scratch_dir,"dem.tif"), overwrite=T)
writeOGR(pnts,paste0(scratch_dir,"."),"gages", driver = "ESRI Shapefile", overwrite_layer = T)

#Fill "single cell" depressions
system(paste(paste(wbt_dir),
             "-r=FillSingleCellPits",
             paste0("--wd=",scratch_dir),
             "--dem='dem.tif'",
             "-o='dem_breach_minor.tif'"))

#Gaussian Filter
system(paste(paste(wbt_dir), 
             "-r=GaussianFilter", 
             paste0("--wd=",scratch_dir),
             "-i='dem_breach_minor.tif'", 
             "-o='dem_filter.tif'",
             "--sigma=3"))

#Breach larger depressions
system(paste(paste(wbt_dir), 
             "-r=BreachDepressions", 
             paste0("--wd=",scratch_dir),
             "--dem='dem_filter.tif'", 
             "-o='dem_breach_major.tif'"))

#Create Flow Accumulation Raster
system(paste(paste(wbt_dir), 
             "-r=D8FlowAccumulation", 
             "--out_type='cells'",
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_major.tif'", 
             "-o='fac.tif'"))

#Create Stream Raster [fac>1000]
fac<-raster(paste0(scratch_dir,"fac.tif"))
fac[fac<threshold]<-NA
fac<-fac*0+1
fac@crs<-p
writeRaster(fac,paste0(scratch_dir,"flowgrid.tiff"), overwrite=T)

#Run flow direction [note we can skip breaching and/or filling sinks b/c we are using NHD data
system(paste(paste(wbt_dir), 
             "-r=D8Pointer", 
             paste0("--wd=",scratch_dir),
             "--dem='dem_breach_major.tif'", 
             "-o='fdr.tif'",
             "--out_type=sca"))

#Create pour pnt raster
system(paste(paste(wbt_dir), 
             "-r=VectorPointsToRaster", 
             paste0("--wd=",scratch_dir),
             "-i='gages.shp'", 
             "--field=ID",
             "-o=pp.tif",
             "--assign=min",
             "--nodata",
             "--base=dem.tif"))

#Jenson Snap Pour point
system(paste(paste(wbt_dir),
             "-r=JensonSnapPourPoints", 
             paste0("--wd=",scratch_dir),
             "--pour_pts='pp.tif'", 
             "--streams='flowgrid.tif'",
             "-o='pp_snap.tif",
             "--snap_dist=1000"
             ))

#Convert back to point file
snapgrid<-raster(paste0(scratch_dir,"pp_snap.tif"))
snappnts<-rasterToPoints(snapgrid, fun=function(x){x>0})
snappnts<-SpatialPointsDataFrame(snappnts[,1:2], data.frame(ID=snappnts))

#Create function to delineate watershed and tabulate LULC
fun<-function(ID){

  #Select Pour Point and Export
  pnt<-snappnts[snappnts$ID.pp_snap==ID,]
  pnt@proj4string<-dem@crs
  writeOGR(pnt,paste0(scratch_dir,"."),"snap", driver = "ESRI Shapefile", overwrite_layer=T)
  
  #Watershed Tool
  system(paste(paste(wbt_dir), 
               "-r=Watershed", 
               paste0("--wd=",scratch_dir),
               "--d8_pntr='fdr.tif'",
               "--pour_pts='snap.shp'", 
               paste0("-o='watershed.tif'")))

  #Import watershed raster into R environment
  ws_grid<-raster(paste0(scratch_dir,"watershed.tif"), overwrite=T)
  
  #Write copy of raster to data directory
  if(!dir.exists(paste0(data_dir,"watershed"))){
    dir.create(paste0(data_dir,"watershed"))
  }
  writeRaster(ws_grid, 
              paste0(data_dir,
                     "watershed/watershed_",
                     pnts@data[pnts$ID==ID,paste(unique_id)], 
                     ".tif"), 
              overwrite=T)
  
  #Lmit LULC grids to watershed
  LULC2001<-LULC2001*ws_grid
  LULC2006<-LULC2006*ws_grid
  LULC2011<-LULC2011*ws_grid
  
  #Calculate Watershed Area
  cell_size<-res(ws_grid)[1]*res(ws_grid)[2]
  area<-cellStats(ws_grid, sum)*cell_size
  
  #Calculate LULC area
  #Create output data.frame
  output<-data.frame(ID=ID,ws_area = area)
  
  #Caluclate water area
  output$urban2001<-length(LULC2001[LULC2001>=20 & LULC2001<30])*cell_size
  output$urban2006<-length(LULC2006[LULC2006>=20 & LULC2006<30])*cell_size
  output$urban2011<-length(LULC2011[LULC2011>=20 & LULC2011<30])*cell_size
  
  #Calculate barren area
  output$barren2001<-length(LULC2001[LULC2001>=30 & LULC2001<40])*cell_size
  output$barren2006<-length(LULC2006[LULC2006>=30 & LULC2006<40])*cell_size
  output$barren2011<-length(LULC2011[LULC2011>=30 & LULC2011<40])*cell_size
  
  #Calculate forest area
  output$forest2001<-length(LULC2001[LULC2001>=40 & LULC2001<50])*cell_size
  output$forest2006<-length(LULC2006[LULC2006>=40 & LULC2006<50])*cell_size
  output$forest2011<-length(LULC2011[LULC2011>=40 & LULC2011<50])*cell_size
  
  #Calculate shrub area
  output$shrub2001<-length(LULC2001[LULC2001>=50 & LULC2001<60])*cell_size
  output$shrub2006<-length(LULC2006[LULC2006>=50 & LULC2006<60])*cell_size
  output$shrub2011<-length(LULC2011[LULC2011>=50 & LULC2011<60])*cell_size
  
  #Calculate Herbaceous area
  output$herb2001<-length(LULC2001[LULC2001>=70 & LULC2001<80])*cell_size
  output$herb2006<-length(LULC2006[LULC2006>=70 & LULC2006<80])*cell_size
  output$herb2011<-length(LULC2011[LULC2011>=70 & LULC2011<80])*cell_size
  
  #Calculate Crop area
  output$crop2001<-length(LULC2001[LULC2001>=80 & LULC2001<90])*cell_size
  output$crop2006<-length(LULC2006[LULC2006>=80 & LULC2006<90])*cell_size
  output$crop2011<-length(LULC2011[LULC2011>=80 & LULC2011<90])*cell_size
  
  #Calculate Wetlands area
  output$wetland2001<-length(LULC2001[LULC2001>=90])*cell_size
  output$wetland2006<-length(LULC2006[LULC2006>=90])*cell_size
  output$wetland2011<-length(LULC2011[LULC2011>=90])*cell_size
  
  #Export output
  output
}

#Run function
output<-lapply(X = snappnts$ID.pp_snap, FUN = fun)
output<-do.call(rbind, output)

#merge output
pnts<-pnts@data
pnts<-merge(pnts, output, by.x="ID", by.y="ID")

#Export output
pnts
}

```


#Step 3: Now lets delineate that watershed

```{r, message=FALSE, warning=FALSE, results='hide'}
#Clean up gages data
gages@data<-gages@data[,c("SOURCE_FEA","DA_SQ_MILE")]

#Select Subbasin
mask<-HUC08
mask<-mask[gages,]

#Create wrapper function
fun<-function(n){
  tryCatch(LULC_WatershedAnalysis(  
              dem=dem,  
              pnts=gages, 
              mask=mask[n,], 
              threshold=1000,
              LULC2001=LULC2001,
              LULC2006=LULC2006,
              LULC2011=LULC2011), 
           error=function(e) c(n, rep(0,24)))}

#Run function
x<-lapply(seq(1,length(mask)), fun)

#pull together data
output<-bind_rows(x)

#Compare area etimates
output<-output[output$DA_SQ_MILE>0,]
output<-output[output$DA_SQ_MILE<2000,]
output$ws_area<-output$ws_area/(1000^2)/(1.60934^2)

#plot
par(mar=c(4,4,1,1))
plot(ws_area~DA_SQ_MILE, data=output, 
     log="xy", xlim=c(0.75,2000), ylim=c(0.75,2000), 
     pch=19, col="grey30", cex=0.5,
     xlab="USGS Estimate [mi^2]", ylab="Jones Estimate [mi^2]", 
     ps=12, cex.lab=14/12, cex.axis=10/12)
abline(a=0,b=1, lwd=2, lty=2, col="dark red")


#Save Image
save.image("backup.R")

```


Okay, so it sorta works. We're biased to "low" estimates where we don't adequately snap to the nearest pour point.  We will manuall check for this. 


Now onto the show

```{r}
#Define pnts
pnts<-read.csv(paste0(data_dir,"input.csv"))
pnts<-SpatialPointsDataFrame(pnts[,c("Longitude","Latitude")], data = data.frame(pnts$FieldActivityId))
pnts@proj4string<-CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")
pnts<-spTransform(pnts, dem@crs)

#Select Subbasin
mask<-HUC08
mask<-mask[pnts,]

fun<-function(n){
  tryCatch(LULC_WatershedAnalysis(  
              dem=dem,  
              pnts=pnts, 
              mask=mask[n,], 
              threshold=1000,
              LULC2001=LULC2001,
              LULC2006=LULC2006,
              LULC2011=LULC2011), 
           error=function(e) c(n, rep(0,24)))}

#Run function
x<-lapply(seq(1,length(mask)), fun)

#pull together data
#output<-bind_rows(x)
#
save.image("initial_output.R")

```

